---
title: "Course Project"
author: "Al Richardson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract


## Introduction

For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

########################################################

## Data Cleaning

1) Convert <i>chr</i> columns to numeric, where possible.
2) 


Remove columns with very sparse data. For example <i>min_yaw_dumbell</i> is a character column, containing numerical data. Approximately 98% of these data are null, i.e."" or "#DIV/0!". It seems that this column is unlikely to contribute much to the model, although that is not at all certain.






########################################################

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
